
---
title: "Bio 110 Week 3"
output:
  pdf_document: default
  html_notebook: default
---

# Lab 3 - Using Digital Microscopy to measure Chlamydomonas flagella


The goals of this tutorial are to:
* learn how to import data in a spreadsheet to R Studio
* calculate the mean of your measured flagella lengths
* calculate the 95% confidence intervals of the mean flagella length
* plot the mean as a point with error bars on a graph

Now that you have collected measurements for 10 Chlamydomonas flagella, you must import your spreadsheet of data to R Studio. The format of the columns of data is important, as well as the titles of the columns of data. 

Before you move on and discover how to load your data into R, it might be useful to go over the following checklist that will make it easier to import the data correctly into R:

With spreadsheets, the first row is usually reserved for the header, while the first column is used to identify the sampling unit;
- Avoid names, values or fields with blank spaces, otherwise each word will be interpreted as a separate variable, resulting in errors that are related to the number of elements per line in your data set;
- If you want to concatenate words, inserting a . in between to words instead of a space;
- Short names are prefered over longer names;
- Try to avoid using names that contain symbols such as ?, $,%, ^, &, *, (, ),-,#, ?,,,<,>, /, |, \, [ ,] ,{, and };
- Delete any comments that you have made in your Excel file to avoid extra columns or NA’s to be added to your file
- Make sure that any missing values in your data set are indicated with NA.


Once you have checked that your spreadsheet follows these basic rules, save it as a .csv file in a BIO 110 folder on the desktop.  

First you must upload your .csv file to R Studio in the bottom right window using the Upload button.  Browse to locate your file and click Upload.  Make sure it's loaded into the bio110-week3 folder.

Before writing any R code, we'll need to load some R packages in order to use some of the specific functions we want. Run the code chunk below to make sure these functions are available to your environement:

```{r}

library(readr)
library(dplyr)
library(Rmisc)
library(ggplot2)
library(rcompanion)


```
 

Now you'll want to load your data as a data frame into R by assigning it to a variable:

NOTE!! data below is just for testing....will remove it (and this sentence) when done. !! END NOTEs

```{r}
# for example:
# myData<-read_csv("myData.csv")
myData<-read_csv("test1.csv")


# confirm by printing your variable
# myData
myData


```





## Generating statistics for plots

Our goal is to generate a plot showing the mean of our data as a point, with error bars displaying a 95% confidence interval. A 95% confidence interval defines a range of values that you can be 95% certain contains the population mean. 

There's a function in R called "summarySE" that can generate the data we need for this plot. Here's the basic structure:

myStats<-summarySE(data=myData, measurevar="MyColumnName", conf.interval = 0.95)

where myData is your data frame, "MyColumnName" is the name of the column to measure, and conf.interval is the percent confidence interval you want to compute. 

Adapt the summarySE command to your data, and run and print it below:

```{r}
# assign a variable to the summarySE function as described above:


# print the variable:

```



This function generates a new data frame with the calculated statistics mean (the column name will match "MyColumnName"), standard deviation (sd), standard error (se), and confidence interval (ci). In order to get and use this data, you'll need to create new variables. You can do this by using the "$" operator, which lets you specify a specific column within a dataframe. For example, if your data frame is named myStats, and you want to get the "ci" data, your variable would look like this:

myCI<-myStats$ci

Below, create variables for the mean, ci, and the upper and lower limits of the confidence interval (hint: the upper limit is the mean+ci, lower is mean-ci)


```{r}
# create variable for the mean

# create variable for the ci value

# create variabel for ci upper limit

# create variabel for ci lower limit


```



Now we have the data needed to create a plot. Beginning with the lab, we're going to be using "ggplot", which is a very popular library for creating data visualizations in R. 


**** MORE GGPLOT STUFF HERE *****



```{r}



ggplot(data=myData, mapping=aes(y=myMean, x="name of group"))+geom_point()+geom_errorbar(mapping=aes(ymin=lower, ymax=upper), width=0.1)+ylim(0,10)+labs(title="my title", x="my x label", y="my y axis label")






```









#That's all I have time for now.  
https://rcompanion.org/handbook/C_03.html
This stats site and its companion R handbook are Greg's favorites.  I thought it might help with finding the code we need.

