
---
title: "Bio 110 Week 3"
output:
  pdf_document: default
  html_notebook: default
---

# Lab 3 - Using Digital Microscopy to measure Chlamydomonas flagella


Might be a good resource or model: http://www.stat.ucla.edu/~rgould/110as02/bsci

The goals of this tutorial are to:
* learn how to import data in a spreadsheet to R Studio
* calculate the mean of your measured flagella lengths
* calculate the 95% confidence intervals of the mean flagella length
* plot the mean as a point with error bars on a graph

Now that you have collected measurements for 10 Chlamydomonas flagella, you must import your spreadsheet of data to R Studio. The format of the columns of data is important, as well as the titles of the columns of data. 

Before you move on and discover how to load your data into R, it might be useful to go over the following checklist that will make it easier to import the data correctly into R:

With spreadsheets, the first row is usually reserved for the header, while the first column is used to identify the sampling unit;
- Avoid names, values or fields with blank spaces, otherwise each word will be interpreted as a separate variable, resulting in errors that are related to the number of elements per line in your data set;
- If you want to concatenate words, inserting a . in between to words instead of a space;
- Short names are prefered over longer names;
- Try to avoid using names that contain symbols such as ?, $,%, ^, &, *, (, ),-,#, ?,,,<,>, /, |, \, [ ,] ,{, and };
- Delete any comments that you have made in your Excel file to avoid extra columns or NA’s to be added to your file
- Make sure that any missing values in your data set are indicated with NA.


Once you have checked that your spreadsheet follows these basic rules, save it as a .csv file in a BIO 110 folder on the desktop.  

First you must upload your .csv file to R Studio in the bottom right window using the Upload button.  Browse to locate your file and click Upload.  Make sure it's loaded into the bio110-week3 folder.

Before writing any R code, we'll need to load some R packages in order to use some of the specific functions we want. Run the code chunk below to make sure these functions are available to your environement:

```{r}

library(readr)
library(Rmisc)
library(ggplot2)
library(rcompanion)


```
 

Now you'll want to load your data as a data frame into R by assigning it to a variable:

NOTE!! data below is just for testing....will remove it (and this sentence) when done. !! END NOTEs

```{r}
# for example:
# myData<-read_csv("myData.csv")
mydata<-read_csv("test1.csv")


# confirm by printing your variable
# myData
mydata


```


Next, you'll want to create a new variable that just has the column data of your measurements. You can do this with this syntax:

myColumn<-datasetVariable$columnName

So if your data frame variable is "myData", and your column header is "Measurements", then it would look like this (the variable name "sampleData" is completely arbitrary):

sampleData<-myData$Measurements

```{r}
# create a new variable with your measurement column data, and print it out to ensure it worked. 

sampledata<-mydata$Measurement

sampledata
str(sampledata)
summary(sampledata)


#just to test!
#ggplot(data=mydata, mapping=aes(x=Measurement))+geom_histogram(bins=15)
# http://rstudio-pubs-static.s3.amazonaws.com/257936_76b37753334b400ea28de4856ccc4f4f.html#calulating_the_confidence_intervals



```






#That's all I have time for now.  
https://rcompanion.org/handbook/C_03.html
This stats site and its companion R handbook are Greg's favorites.  I thought it might help with finding the code we need.

